White Dwarf Mergers on Adaptive Meshes\\ II. Collisions

### 2016-07-15 papers/wdmerger_II/paper.tex
Update the gravitational wave signal

Over the past few years a number of groups have performed simulations of WD collisions
to understand whether they may yield astrophysical transients that look like SNe Ia
\citep{rosswog:2009,raskin:2010,loren-aguilar:2010,hawley:2012,garcia-senz:2013,
kushnir:2013,papish:2015,holcomb:2015}. Head-on collisions rapidly convert a significant
amount of kinetic energy into thermal energy in a small region and thus set up the
conditions for a thermonuclear detonation, and these previous simulations have indeed
found that detonations occur and convert a large amount of carbon/oxygen material into
iron-group elements. These papers varied significantly in the hydrodynamic methods used
(Lagrangian versus Eulerian methods), the methodology used for coupling nuclear
reactions to the hydrodynamics (including variation in the number of isotopes in
the nuclear network and in the evolution of the temperature), and the temporal and
spatial resolution. Consequently they yielded estimates of production of nickel-group
elements that varied significantly, leaving much uncertainty about how such an event
would appear observationally and whether it bears any resemblance to a SN Ia.

@@

Paper I established our methodology for calculating the gravitational wave signature
of a white dwarf merger using the post-Newtonian framework of \citet{blanchet:1990}.
While a typical merger event has a characteristic profile including a chirp and ringdown,
white dwarf collisions feature sharply different gravitational wave signatures, and
hence may be interesting targets for observation for future low frequency gravitational
wave detectors \citep{eLISA}. The main caveat for these events compared to events
such as neutron star mergers is that due to the event being inherently less energetic,
the typical distance scale on which we could observe a white dwarf merger or collision
is the size of our own galaxy, with extragalactic events being unlikely to be
detected, at least with the first generation of space-based gravitational wave
detectors. \autoref{fig:gw_signal} is a plot of the gravitational wave strain for
the representative case of \autoref{sec:parameters}. We use the convention that the
absolute scale of the strain is normalized to a distance of 10 kpc (the strain scales
inversely with the source distance). The shape of the profile agrees qualitatively
with the gravitational wave signature predicted by \cite{loren-aguilar:2009:collisions},
though because they used an unequal mass pair the results do not match exactly.

\begin{figure}
  \centering
  \includegraphics[scale=0.425]{plots/gw_signal}
  \caption{Gravitational wave strain for the head-on collision of two $0.64\ \msolar$
           white dwarfs. The strain is normalized to a source distance of 10 kpc.
           \label{fig:gw_signal}}
\end{figure}

>>>>

Over the past few years a number of groups have performed simulations of WD collisions
to understand whether they may yield astrophysical transients that look like SNe Ia
\citep{rosswog:2009,raskin:2010,loren-aguilar:2010,hawley:2012,garcia-senz:2013,
kushnir:2013,papish:2015,holcomb:2015}. Head-on collisions rapidly convert a significant
amount of kinetic energy into thermal energy in a small region and thus set up the
conditions for a thermonuclear detonation, and these previous simulations have indeed
found that detonations occur and convert a large amount of carbon/oxygen material into
iron-group elements. These papers varied significantly in the hydrodynamic methods used
(Lagrangian versus Eulerian methods), the methodology used for coupling nuclear
reactions to the hydrodynamics (including variation in the number of isotopes in
the nuclear network and in the evolution of the temperature), and the temporal and
spatial resolution. Consequently they yielded estimates of production of nickel-group
elements that varied significantly, leaving much uncertainty about how such an event
would appear observationally and whether it bears any resemblance to a SN Ia.
See Table 4 of \cite{garcia-senz:2013} for a summary of the outcome of many of these studies.

@@

Paper I established our methodology for calculating the gravitational wave signature
of a white dwarf merger using the post-Newtonian framework of \citet{blanchet:1990}.
While a typical merger event has a characteristic profile including a chirp and ringdown,
white dwarf collisions are noticeably distinct in their gravitational wave signatures, and
hence may be interesting targets for observation for future low frequency gravitational
wave detectors \citep{eLISA}. The strain increases monotonically for a collision until impact,
instead of oscillating back and forth at a characteristic frequency as for mergers.
The main caveat for these events compared to events
such as neutron star mergers is that due to the event being inherently less energetic,
the typical distance scale on which we could observe a white dwarf merger or collision
is the size of our own galaxy, with extragalactic events being unlikely to be
detected, at least with the first generation of space-based gravitational wave
detectors. \autoref{fig:gw_signal} is a plot of the gravitational wave strain for
the representative case of \autoref{sec:parameters}, where we disabled the normal stopping
criterion and let the simulatin run until $t = 20$. We use the convention that the
absolute scale of the strain is normalized to a distance of 10 kpc (the strain scales
inversely with the source distance). The shape of the profile agrees qualitatively
with the gravitational wave signatures predicted by \cite{loren-aguilar:2009:collisions}
and \cite{garcia-senz:2013}.

\begin{figure}
  \centering
  \includegraphics[scale=0.425]{plots/gw_signal}
  \caption{Gravitational wave strain $h_+$ for the head-on collision of two $0.64\ \msolar$
           white dwarfs. The strain is normalized to a source distance of 10 kpc. We do not
           show the other polarization, $h_\times$, as for an equal-mass head-on collision it vanishes
           identically.
           \label{fig:gw_signal}}
\end{figure}
